---
title: "HPM R learners - Summer 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::read_chunk("../code/r-users-scheduling.r")
```

Last Updated `r format(Sys.Date(), format="%m/%d/%Y")`

# {.tabset}

This will be the R learners website for the summer! Most all of my resources will be posted here. 

Contact: Natalie Smith [natsmith@live.unc.edu](natsmith@live.unc.edu)

## Intro

We will meet regularly on Mondays from 11am-1pm. I'll spend 1 hour (max) reviewing the topic for that day and the different packages, etc. that are involved. The second hour will be working through examples independently (but still together) and is totally optional -- you can feel free to work through the examples on your own whenever works for you. 

This will be very casual -- kids, pets, eating are all okay! Interupt me anytime via chat or audio to ask questions and *please* tell me to slow down if necessary. 

First month of topics: 

+ June 1: Workflows in R - projects, scripts, and markdown files. Integration with GitHub -- what is git & why you should use it, and using GitHub pages to make easy-to-access html webpages for your projects!
+ June 8: What are R objects, and why do I have to care? Types of variables in R - factors! Begin data management in R -- using the `tidyverse`
+ June 15: Natalie unavailable -- take week off, or someone else lead a session? Descriptive statistics in R? Can share my code for this and people work independently, or someone else moderate? 
+ June 22: Begin plotting!
+ June 29: Continue plotting! 




#### Potential topics

+ Data management with the `tidyverse` package(s). Would probably suggest two weeks of this.
    + Types of variables in R -- factors are the most unique to R specifically
    + Merging, subsetting, and transposing data
    + Creating new variables
+ Working with objects in R - what are they, how, and why??? `r emo::ji("angry")`
+ Descriptive statistics in R
    + There are many, many packages - how to choose? 
    + Overview with `skimr`
    + Missing data with `naniar` package
    + Full tables with `tableone`
    + There's also `stargazer` and some others
+ Basic modeling - lm, logistic, others?
+ Other stat topics? 
+ Plotting with R - endless `ggplot2` options here! Can review all sorts of x-y plots, there are some new options for making maps. 
+ R projects, R markdown, GitHub, & GitHub pages (how I am making this website, very nice for collabs)


#### Meeting times

Some fun data management to wrangle our data into shape for graphing
```{r dm, message=FALSE, warning=FALSE}
```

Now let's make a beautiful graph
```{r graph}
```





## Getting started

Two general options for how to access and use R

1. Use R through R studio on your actual computer
    + Download R v4.0.0 [here](http://archive.linux.duke.edu/cran/)
    + Download R studio [here](https://rstudio.com/products/rstudio/download/#download)

2. Use [rstudio.cloud](rstudio.cloud) without the commitment of installing on your computer. *disclaimer: I don't know as much / anything about this*
    + Don't have to download R software and the R studio wrapper
    + Possibly less functional for all the bells and whistles -- not sure about GitHub integration, etc. But probably just fine for learning the basics, and much less potential annoyance with installations, etc. 

3. *Optional* Make a [GitHub](https://github.com/) account
    + Not necessary to use R. Is necessary to make a website (like this one) and use GitHub for collaborations, etc. 
    
4. *Optional* Download [GitHub Desktop](https://desktop.github.com/). This is an interface that makes git not-scary and allows you to easily commit, push, and pull from your online GitHub account. 


## Sessions {.tabset}

### June 8 - Workflows

Specific things to cover:
+ R projects. Relative file paths, folder structure, etc. 
+ R scripts. Installing packages. Loading packages. File names, integrating with Markdown. 
+ R markdown. First steps. Knitting to pdf vs. html. Chunks. Pros and cons of code-in-scripts vs. code-in-markdown. 
+ GitHub integration. Integrating repositories with R projects, gitignore files, committing, pushing, pulling. 
+ Making project website with GitHub pages. 

Exercises: 
+ Clone an example repository [add link here](github.com)
+ Pull 
+ Make a change to the markdown file with your name and email address.
+ Push
+ Did the website update?





## Resources

This is my own compilation of useful R tidbits that don't have a good place to live, and my go-to resources for various things in R. 


+ Tools > Global Options > General > Make sure the following things are done. Best practice for reproducible work!
    + 'restore .Rdata' is unchecked
    + saving .Rdata is set to 'Never'
    + 'always save history' is unchecked. 
+ Emojis in R markdown
    + Install by: `devtools::install_github("hadley/emo")` (need `devtools` package installed)
    + [emoji package](https://github.com/hadley/emo)
    + [Emoji search](https://emoji.muan.co/)
+ [Google it](https://twitter.com/andrewheiss/status/1263652454329208839)
+ Online resources
    + Intro to `naniar` package [here](https://allisonhorst.shinyapps.io/missingexplorer/#section-introduction)
    + *Very good* `ggplot` intro by [Andrew Heiss](https://twitter.com/andrewheiss) [here](https://datavizm20.classes.andrewheiss.com/). Also some `tidyverse` but not enough to fully understand it
    + R studio cheatsheets [here](https://rstudio.com/resources/cheatsheets/)
+ Guide for good coding in social sciences, from some economists [here](https://web.stanford.edu/~gentzkow/research/CodeAndData.pdf). R is great because you can very easily do everything they suggest, and document with Rmd, and have more than one dataset open at a time `r emo::ji("fire")`
    + automation: functions are very easy in R
    + version control: git integrates well with R
    + directories: R projects
    + keys: not specific to software (keep data stored with identifiers)
    + abstraction: functions round 2
    + documentation: R markdown, easier options than local variables. 
    
    

    
    
    